Symmetric Peer-to-Peer Applications

### 2022-09-11 tpd-22.tex
...

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Nodes in the application form a dynamic graph network and communicate only
with direct neighbours.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware that runs on each peer to
ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate its whole
execution from the beginning (or from a snapshot in between).
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time.
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
When a peer leaves the network, nothing needs to be done.
The middleware just needs to ensure that a peer that joins the network receives
all events in order, which is trivial since they all hold an absolute
timestamp.
\end{enumerate}

@@

Figure~\ref{TODO} results...

We now detail how we instrument the middleware implementation measure the
properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
\item \textbf{Real-time pace of peers:}
We start each peer with a random delay of up to $10s$ to force a time mismatch.
Each peer outputs its current tick continuously, allowing us to compare them
in the shared timeline.
Each time we see a greater tick, we count how many smaller ticks from other
peers appear in the timeline, which means that they are ...
\item \textbf{Correctness of unstable peers:}
- state (even for stable peers)
- time
%
\end{enumerate}

@@

>>>>

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic graph network and communicate only
with direct neighbours, as a typical unstructured peer-to-peer network.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware that runs on each peer to
ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate its whole
execution from the beginning (or from a snapshot in between).
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time (e.g., from rollbacks or timer
        polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware just needs to ensure
that it receives all events in order, which is trivial since they all hold an
absolute timestamp.
\end{enumerate}

@@

Figure~\ref{TODO} results...
- colors indicate feasibility
%
We now detail how we instrument the middleware implementation to measure the
properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
\item \textbf{Real-time pace of peers:}
Every time a peer receives \code{SYNC} packets and is behind, we output its id
and how much ticks it needs to fast forward.
At the end, we measure the percentage of the sum of all forward ticks over the
total simulation ticks (\code{TODO} in Figure~\ref{TODO}).
As an example, if all peers sum $1000$ forward ticks over $100k$ accumulated
simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a time
mismatch.
%
Each peer outputs its current tick continuously, allowing us to compare them
in the shared timeline.
Each time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers (considering latency), which would constitute
time violations (\code{TODO} in Figure~\ref{TODO}).
We expect this counter to be close to zero considering the whole execution.
\item \textbf{Correctness of unstable peers:}
- state (even for stable peers)
- time
%
\end{enumerate}

@@

- check if all ticks of all peers are in the same state
